Approximately Exact Calculations for Linear Mixed Models

### 2015-05-30 ms.tex
HMO Bayes

\subsubsection{A log RL Analysis}
For a $\logRLssorig$ analysis there are 43 lines, as shown in Figure~\ref{fig:hmolines}.
Running the algorithm on the box $B_1$ determined by the lines' maximum intercepts on the $\sigesq$ and $\sigssq$ axes with the settings
\begin{equation*}
	\maxit=15; \hspace{1cm} \epsilon=1; \hspace{1cm}
	\delta_e=\log(10); \hspace{1cm} \delta_s=\log(10); \hspace{1cm} \text{ratio $=$ TRUE}; \hspace{1cm} M=5
\end{equation*}
results in the the output displayed in Table~\ref{table:hmo_HH11_run}, which shows the state of the algorithm after iterations 1 through 15: the numbers of active and inactive boxes and the current value of the lower bound $L$ on $\max\log f$.  We see that boxes are steadily transferred from the active to the inactive list and that $L$ increases monotonically.  After 15 iterations there is a total of 895 boxes, which are displayed in Figure~\ref{fig:hmo1}.

@@

For a Bayesian analysis there are 44 lines, as shown in Figure~\ref{fig:hmoBayeslines}.  Figure~\ref{fig:hmoBayeslines} differs from Figure~\ref{fig:hmolines} in that it includes a horizontal line for $j = s_z+2$ and the position of the vertical line for $j = s_z+1$ is slightly shifted.  Because there is only a slight change from the $\logRLorig$ analysis, we began by running the algorithm with the refined control parameters from the $\logRLorig$ analysis.  After 15 iterations there were still active boxes, so we ran the algorithm again with the same control parameters except $\maxit=20$.  Table~\ref{table:hmo_HH11Bayes} shows the output.  After 20 iterations there are over 3,000,000 boxes, over 2,000,000 of which are still active.
\begin{figure}
	\centering
	\includegraphics[width=.45\linewidth]{figs/hmolines_HH11_Bayes.pdf}
	\caption{The 44 lines, $j=1, 2, \dots, 44$, for the Bayesian analysis of the HMO data.
	              Lines 43 and 44 are dashed.}
	\label{fig:hmoBayeslines}
\end{figure}

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
Iteration & $n$ active boxes & $n$ inactive boxes & $L$\\
1 & 4 & 0 & $-1350.29$\\
2 & 16 & 0 & -1303.20\\
3 & 64 & 0 & -1279.93\\
4 & 256 & 0 & -1270.26\\
5 & 1024 & 0 & -1265.89\\
6 & 4060 & 9 & -1263.71\\
7 & 13740 & 634 & -1262.60\\
8 & 14904 & 10648 & -1262.01\\
9 & 10180 & 23007 & -1261.71\\
10 & 7364 & 31346 & -1261.71\\
11 & 6012 & 37207 & -1261.71\\
12 & 8192 & 41171 & -1261.71\\
13 & 16384 & 45267 & -1261.71\\
14 & 32768 & 53459 & -1261.71\\
15 & 65536 & 69843 & -1261.71\\
16 & 131072 & 102611 & -1261.71\\
17 & 262144 & 168147 & -1261.71\\
18 & 524288 & 299219 & -1261.71\\
19 & 1048576 & 561363 & -1261.71\\
20 & 2097152 & 1085651 & -1261.71\\
\hline
\end{tabular}
\caption{The state of the algorithm after each of 20 iterations for the Bayesian analysis of the HMO data.}
\label{table:hmo_HH11Bayes}
\end{table}

Figure~\ref{fig:hmoBayes} shows the boxes and $\log \pi(\sigesq,\sigssq)$.  Figure~\ref{fig:hmoBayes} is similar to Figure~\ref{fig:hmo2} except for the addition of the outlined blue boxes near $\sigssq = 0.04$.  In fact, those boxes all have lower boundaries $\sigssq = 0.04673004$, upper boundaries $\sigssq = 0.04768372$, $L^b \in (-1270.00, -1266.09)$, $U^b \in (-1252.877, -1248.524)$, and $U^b-L^b \in (17.12299, 18.01643)$, whereas the red dot shows where $\max L^b = -1261.705$ occurs.  The blue boxes near $\sigssq = 0.04$ are introduced by the $\IG(1.1,0.1)$ prior, which has a mean of 1, an infinite variance, and a peak at 0.04761905.  The usual analyses of the posterior density are carried out either by MCMC --- in which case one hopes that one's Markov chain happens to sample all regions of high density --- or by evaluating the posterior on a grid --- in which case one hopes one's grid is fine enough to capture all regions of high density.  With our analysis there is no need to hope; the algorithm is certain to find all regions of high density.

\begin{figure}
  \begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=.8\linewidth]{figs/hmo_HH11Bayes_boxes.jpg}
	\caption{Locations of the boxes.}
	\label{fig:hmoBayesboxes}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=.8\linewidth]{figs/hmo_HH11Bayes_rll.jpg}
	\caption{Grayscale shows $L^b$ in each box.  Red dot shows $b^* \equiv\argmax L^b$.
	Boxes outlined in blue have $U^b \ge L^{b^*}$.  Axes are logarithmic.}
	\label{fig:hmoBayesrll}
  \end{subfigure}
  \caption{The 3,182,803 boxes in the Bayesian analysis of the HMO data.
               $\maxit=20; \epsilon=1; \delta_e=0; \delta_s=0; \text{and\ } M=10$.}
  \label{fig:hmoBayes}
\end{figure}

>>>>

\subsubsection{A log RL Analysis}
For a $\logRLssorig$ analysis there are 43 lines, as shown in Figure~\ref{fig:hmolines}.
Running the algorithm on the box $B_1$ determined by the lines' maximum intercepts on the $\sigesq$ and $\sigssq$ axes with the settings
\begin{equation*}
	\maxit=\infty; \hspace{1cm} \epsilon=1; \hspace{1cm}
	\delta_e=0; \hspace{1cm} \delta_s=0; \hspace{1cm} M=7
\end{equation*}
results in the the output displayed in Table~\ref{table:hmo_HH11_run}, which shows the state of the algorithm after iterations 1 through 15: the numbers of active and inactive boxes and the current value of the lower bound $L$ on $\max\log f$.  We see that boxes are steadily transferred from the active to the inactive list and that $L$ increases monotonically.  After 15 iterations there is a total of 895 boxes, which are displayed in Figure~\ref{fig:hmo1}.

@@

For a Bayesian analysis there are 44 lines, as shown in Figure~\ref{fig:hmoBayeslines}.  Figure~\ref{fig:hmoBayeslines} differs from Figure~\ref{fig:hmolines} in that it includes a horizontal line for $j = s_z+2$ and the position of the vertical line for $j = s_z+1$ is slightly shifted.
\begin{figure}
	\centering
	\includegraphics[width=.45\linewidth]{figs/hmolines_HH11_Bayes.pdf}
	\caption{The 44 lines, $j=1, 2, \dots, 44$, for the Bayesian analysis of the HMO data.
	              Lines 43 and 44 are dashed.}
	\label{fig:hmoBayeslines}
\end{figure}
Because the $\logRLorig$ analysis in Figure~\ref{fig:hmorll} shows low values for $\sigesq, \sigssq > 1000$, we run the Bayesian analysis on the box $B^0 = (0,1000) \times (0,1000)$, again with the control parameters
\begin{equation*}
	\maxit=\infty; \hspace{1cm} \epsilon=1; \hspace{1cm}
	\delta_e=0; \hspace{1cm} \delta_s=0; \hspace{1cm} M=7.
\end{equation*}
The algorithm finished in 22 iterations.  Table~\ref{table:hmo_HH11Bayes} shows the output.
\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
Iteration & $n$ active boxes & $n$ inactive boxes & $L$\\
1 & 4 & 0 & $-\infty$\\
2 & 16 & 0 & -1313.42\\
3 & 52 & 3 & -1285.05\\
4 & 132 & 22 & -1270.76\\
5 & 264 & 88 & -1263.98\\
6 & 628 & 195 & -1260.47\\
7 & 1780 & 378 & -1258.71\\
8 & 4188 & 1111 & -1257.70\\
9 & 5676 & 3880 & -1256.88\\
10 & 5136 & 8272 & -1255.56\\
11 & 5528 & 12026 & -1254.20\\
12 & 8236 & 15495 & -1252.84\\
13 & 12488 & 20609 & -1251.50\\
14 & 23852 & 27134 & -1250.26\\
15 & 22740 & 45301 & -1249.22\\
16 & 43228 & 57234 & -1248.77\\
17 & 160208 & 60410 & -1248.77\\
18 & 331900 & 137643 & -1248.77\\
19 & 575196 & 325744 & -1248.77\\
20 & 942860 & 665225 & -1248.77\\
21 & 722412 & 1427482 & -1248.77\\
22 & 0 & 2149894 & -1248.77\\
\hline
\end{tabular}
\caption{The state of the algorithm after each of 22 iterations for the Bayesian analysis of the HMO data.}
\label{table:hmo_HH11Bayes}
\end{table}
Figure~\ref{fig:hmoBayes} is similar to Figure~\ref{fig:hmo2} except for the addition of the outlined blue boxes near $\sigssq = 0.04$.  In fact, those boxes all have lower boundaries $\sigssq = 0.04673004$, upper boundaries $\sigssq = 0.04768372$, $L^b \in (-1270.00, -1266.09)$, $U^b \in (-1252.877, -1248.524)$, and $U^b-L^b \in (17.12299, 18.01643)$, whereas the red dot shows where $\max L^b = -1261.705$ occurs.  The blue boxes near $\sigssq = 0.04$ are introduced by the $\IG(1.1,0.1)$ prior, which has a mean of 1, an infinite variance, and a peak at 0.04761905.  The usual analyses of the posterior density are carried out either by MCMC --- in which case one hopes that one's Markov chain happens to sample all regions of high density --- or by evaluating the posterior on a grid --- in which case one hopes one's grid is fine enough to capture all regions of high density.  With our analysis there is no need to hope; the algorithm is certain to find all regions of high density.

\begin{figure}
  \begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=.8\linewidth]{figs/hmo_HH11Bayes_boxes.jpg}
	\caption{Locations of the boxes.}
	\label{fig:hmoBayesboxes}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=.8\linewidth]{figs/hmo_HH11Bayes_rll.jpg}
	\caption{Grayscale shows $L^b$ in each box.  Red dot shows $b^* \equiv\argmax L^b$.
	Boxes outlined in blue have $U^b \ge L^{b^*}$.  Axes are logarithmic.}
	\label{fig:hmoBayesrll}
  \end{subfigure}
  \caption{The 2,149,894 boxes in the Bayesian analysis of the HMO data.}
  \label{fig:hmoBayes}
\end{figure}
